### 2022-03-26 ms/PrestigeBias.tex
update

\section{Discussion}
Cultural transmission is the phenomenon of which cultural elements, in the form of attitudes, values, beliefs, and behavioral patterns, are transmitted between individuals, typically via copying.
Some cultural traits can be more likely to be copied by others, regardless of their frequency in the population.  
Such transmission biases are common in cultural transmission processes. 
Many models are based on the assumption that success can be correctly identified, and easily copied.
Here we assume that success isn't correctly identified, therefore individuals may use other indicators to try and estimate the success of potential role-models.
We believe, as \citet{complexityPaper} suggest, that \textit{prestige biases} are more common in nature than success biases, since estimating success is harder.
We believe prestige is composed of two main components: a trait that indicates success (but doesn't guarantee it), and the influence the individual already has on others, i.e number of individuals already chose him as a role-model.
We suggest a model for \textit{prestige bias}, inspired by the model \citet{evolutionBook} have suggested, and added the \textit{influence bias} to it.
We approximated our models using various distributions, and compared them to the original model using simulations.
We showed that a \textit{Rich getting richer} type of model can be approximated well by the general binomial distribution and the dirichlet multinomial distribution.
We experimented with constant and changing environment in our model, and created a variation of a binary model for easier mathematical and computational analysis.
We believe that in this era of social media it is easy to estimate one's influence over others.
It is therefore crucial to model the cultural biases more realistically than success bias based model, and we believe including influence is crucial for that purpose.\\
With a more realistic model of a common cultural transmission bias, we may be able to better understand decision-making processes in humans, including life-changing choices such as occupation or a life partner. 
Our model can be expanded in multiple ways: observing the effects of different bias functions, including errors in estimating the influence, combining factors of cost when copying from an influential role model (not all could afford to copy from the most popular role-model), and analyzing the differences when including several optimal values for the indicator trait (multiple preference traits in the population).

\clearpage
\section{Summaries}
\subsection{Animal leadership \citep{animal_leadership}}
There are two main approaches to decision makings of groups in nature: leadership and consensus. Prey leaders would lead the pact when traveling, while other animal group leaders will decide on a nesting site or foraging patches. They found out that leadership is observed mostly when there is a profound social network in the group, and when there are individuals that present leadership behavior. Leaders would usually be high ranking members in the group, such as elders, individuals with many kin relations, or posses other dominant traits. When no individual posses such traits, or when the social network is lacking, a consensus is more likely to occur.
When a leader is present, they will have greater selection costs, such as higher risk for predation, being poisoned by unknown experimental patch, but also greater benefits. For example, given the route to the foraging site was successful, the leader and his closest followers would gain most of the food, unlike in a consensus, where the food would be shared more equally.
It appears leaders appear in simple organisms as well, like fish. In these organisms however, the leader would usually be the hungriest or the weakest, while the rest would prefer to follow, minimizing their costs. In baboons however, \citet{animal_leadership} describe many benefits for the closest associates of the dominant male, such as protection from predators. This is an instance of sexual-selection, where the leader, whose survival chances are lower, gains more sexual partners due to the benefits. ("The greater the risk, the greater the reward")
In humans, leadership also has its perks and costs. Leaders can make decisions that would benefit them and their closest followers the most, while still maintaining group cohesion. However, wrong decision making that would harm the group could result in harm (media, social status, even violent behavior of subjects on certain cases).

\subsection{Prestige as a type of leadership \citep{dual_leadership}}
In their paper, \citet{dual_leadership} suggest a different view of leadership. They say that most discussions assume there is one type of leadership, and so they differ in their definition. \citet{dual_leadership} suggest a way to solve said contradiction by defining two types of leaderships: prestige-based and dominance-based.
They present classical views of leaderships by Confucius and Machiavelli. Confucius views leaders as role models who exercise influence through possessing superior knowledge, skills,
and (outstanding) personal qualities. This description is very similar to our indicator trait. By contrast, Machiavelli views leaders as rulers who exercise influence by imposing costs through (the threat of) punishment. They say that these two opposing views are both partially supported by the available evidence but each one on its own offers an incomplete view into the complex and dynamic processes of leadership.

Our current model doesn't reflect the model described in this article, but several adjustments could be made in order to match it. If we assume there's a correlation between trait value to a type of leadership (so in our binary model, one trait would be of prestige, and the other of dominance) we can implement their suggested model. For that we would need to add cost-benefit parameters, so the ones choosing prestige will be rewarded, but pay more, while the ones choosing dominance would pay less, but gain less benefits. It could be interesting to see the dynamics and relations between our model parameters and these cost-benefit parameters.

\subsection{Evidence from Fijian villages for adaptive learning biases \citep{fijian_social_bias}}
\citet{fijian_social_bias} researched Fijian villages, looking for evidence of social learning biases and their origins. They mention that:
\begin{quote}
evolutionary theorists propose that natural selection has favored the emergence of psychological biases for learning from those individuals most likely to possess adaptive information.
\end{quote}
Their goal is to bridge from the laboratory to the field by examining if and how these biases emerge in a small-scale society.
During their research they found that:
\begin{quote}
Fijian villagers (ages 10 and up) are biased to learn from others perceived as more successful/knowledgeable, both within and across domains (prestige effects).
\end{quote}
Their research shows promising evidence for our prestige model, suggesting that copying from others who are \textit{perceived} as successful, rather than actually are successful. In their paper, they show that the social networks representing copier-role-model relationships are centralized, suggesting:
\begin{quote}
This degree of centralization is consistent with the prediction that people substantially share notions about who is a good cultural model (network centrality), but that individuals’ model selections are influenced by multiple factors.
\end{quote}
We see here support for both our indicator trait and our influence bias in their data.

\subsection{Using social influence to reverse harmful traditions \citep{harmful_traditions}}
Social influence could be a powerful tool to end traditions, harmful or advantageous. Harmful traditions can be child marriage, open defecation, and domestic violence, to name a few. 
\citet{harmful_traditions} suggest a mechanism called \textit{spillover}. By their definition, a spillover is when an intervention affect a large enough group in a target population, so that others not included in the intervention starts changing their behavior as well. In their research, they found that there are individuals who act as \textit{agents}, who are often looked upon, and therefore they are ideal targets for interventions. This is the same concept as our role-models, where a more prestigious individual will be copied more, therefore spreading his trait wider in the population. Their research support therefore in our assumption that there are social biases, conformist influence specifically. They also suggest a way to use this phenomena to change existing traditions in a population. It is very clear however, that just as it can be used to end harmful traditions, the same agents could be used for any negative way that comes to mind.

\subsection{The evolution of prestige \citep{prestige_evolution}}
One cannot mention prestige and cultural transmission without the work of \citet{prestige_evolution}. They as well distinguish between prestige and dominance. The latter is is defined by acquiring social status by using aggression, intimidation and violence. They also mention dominance is more common than prestige in non humans. Their definition of prestige is somewhat synonymous with ours. Specifically prestige is composed both of estimation in the eyes of people (our indicator/success trait) and commanding position in people's minds, i.e number of copiers people think they have, which they define as \textit{influence} (similar to our definition for influence).
In their paper, they show that prestige evolved from natural selection, as an efficient process to extract reproductive benefit from the flow of socially transmitted information. Simply put, prestige is a natural step where social learning exists, due to saving costs of individual learning.
It could be interesting in the future to enhance our model with this idea, i.e observing the copier trait of \textit{evaluation}, rather than only observing the evolution of the indicator trait copied.
\citet{prestige_evolution} suggest that the most skilled role-models will, on average, end up with most copiers. This is consistent with our findings, and in a way our model is a mathematical equivalent to their suggested model.

\subsection{Biased learning may be disfavored when payoffs are uncertain \citep{best_of_k}}
\citet{best_of_k} remind that social learning not only takes the form of random copying of other individuals, but also involves learners’ choice of what to learn or from whom to learn. They suggest a best-of-K model where an individual samples k role-models and choose the one he deems most "successful". They mentioned that a previous mathematical analysis has shown that it may sometimes result in maladaptive cultural evolution when the payoffs associated with cultural variants vary stochastically. In such a case, learners may be selectively disfavored and in the long run replaced by unbiased learners, who simply copy someone chosen at random. They develop new mathematical models that are more simplified and mathematically tractable. They found that best-of-k learning, unlike unbiased learning, can facilitate the invasion of an on average inferior variant that sometimes gives a very high payoff. Our model, which includes influence bias, is consistent with that claim. When a maladaptive trait is "piggybacking" a role-model with high influence, said trait could spread in the population, as mentioned. In addition, they show that best-of-k learning can be stable against invasion by unbiased learning if social learning is sometimes combined with individual learning. Our model is based on copying based learning only, but it could be interesting to combine it with individual learning and see how it affects the dynamics of the population.

\subsection{Prestige in medicine \citep{medical_prestige}}
So far we presented the existence of prestige in nature and in humans. We mostly discussed it in regard of social ranking in animals, and what are the costs and benefits. In humans, we mostly addressed spread of adaptive or maladaptive traits due to the prestige bias. \citet{medical_prestige} present a specific and important effect of prestige - its significance for medical specialties and diseases.
They examined literature from 1950 to 2005 regarding the effects of prestige on medicinal practices. They discovered that active, specialized, biomedical, and high-technological types of medicine on organs in the upper part of the bodies of young and middle-aged people were accorded high levels of prestige, while medicine with opposite characteristics had low levels of prestige. They have concluded that such differences in prestige bear consequences for actual priority setting in healthcare systems. They discovered that surgery counts as the most prestigious specialty, while psychiatry is the less prestigious. In addition, doctors tend to rank practices that require more time to master as more prestigious, while other procedures that are considered \textit{easier} are less prestigious. \citet{medical_prestige} found that the advance in technology and research was in accordance with the prestige rankings.

\subsection{Collective brains \citep{collective_brains}}
We showed cultural transmission is a process that manifests in many species, with emphasis on humans. We also displayed similarities between this process and genetic transmission, while presenting differences between them, specifically selective biases such as influence and prestige. \citet{collective_brains} offer a take on prestige as a factor of human physical evolution. They present a concept called \textit{cultural brains} - brains that evolved primarily for the acquisition of adaptive knowledge.
They build on the hypothesis of \citet{social_brains} that shows that larger, more complex brains can store and manage more information and in turn, this information can support the costs of a larger brain.
\citet{collective_brains} built a model that predicts a strong relationship between brain size and group size, because group size also provides access to more adaptive knowledge. They later present their \textit{cumulative cultural brain} hypothesis, an approach which proposes that human brains have evolved with an ability and proclivity for selective, high fidelity social learning. As part of this process, there are a variety of strategies and biases that have evolved to hone in on the most adaptive knowledge. These strategies and biases include direct and indirect cues of the popularity of cultural traits (e.g. success and prestige biases).
In short, they suggest that some of the reasons for the extreme increase in brain size in humans, are the ability to "cheaply" acquire adaptive knowledge, i.e transmission biases, such as prestige.

\subsection{Prestige-biased cultural learning \citep{prestige_cultural_learning}}
Culture–gene coevolutionary (CGC) researchers theorize that humans should possess several learning biases which robustly enhance the fitness of cultural learners. 
\citet{prestige_cultural_learning} report the first direct tests in children of CGC's prediction of \textit{prestige bias}, a tendency to learn from individuals to whom others have preferentially attended, learned or deferred.
Their definition of prestige is similar to our \textit{influence bias}, and brings concrete proof of its existence and effects. Their study showed that the odds of 3-4 years-old children learning from an adult model to whom bystanders had previously preferentially attended for 10 seconds were over twice those of their learning from a model whom bystanders ignored.
In addition to this first study, they also discovered prestige effects are domain-sensitive. They saw that prestigious models were listened to by most when demonstrating artifact-use, but not as much as when presenting food preferences. It lead \citet{prestige_cultural_learning} to believe that when the trait is costly to learn individually, prestige will have a higher effect. It would be interesting to include costs in our model to try and observe these effects and their dynamics in the simulations of a larger population than this study.

>>>>

\section{Discussion}
\subsection{summary}
Cultural transmission is the phenomenon of which cultural elements, in the form of attitudes, values, beliefs, and behavioral patterns, are transmitted between individuals, typically via copying.
Some cultural traits can be more likely to be copied by others, regardless of their frequency in the population.  
Such transmission biases are common in cultural transmission processes. 
Many models are based on the assumption that success can be correctly identified, and easily copied.
Here we assume that success isn't correctly identified, therefore individuals may use other indicators to try and estimate the success of potential role-models.
We believe, as \citet{complexityPaper} suggest, that \textit{prestige biases} are more common in nature than success biases, since estimating success accurately is harder.
\textbf{We investigated the effects of prestige on a population:} we studied the behavior of an invading trait, analyzed its dynamics mathematically, and extended the basic constant environment to a changing one.
We believe prestige is composed of two main components: a trait that indicates success (but doesn't guarantee it), and the influence the individual already has on others, i.e number of individuals already chose him as a role-model.
We suggest a model for \textit{prestige bias}, inspired by the model \citet{evolutionBook} have suggested, and added the \textit{influence bias} to it.
\textbf{We approximated our models using various distributions, and compared them to the original model using simulations.}
We showed that a \textit{Rich getting richer} type of model can be approximated well by the general binomial distribution and the dirichlet multinomial distribution.
We experimented with constant and changing environment in our model, and created a variation of a binary model for easier mathematical and computational analysis.
We believe that in this era of social media it is easy to estimate one's influence over others.
It is therefore crucial to model the cultural biases more realistically than success bias based model, and we believe including influence is crucial for that purpose.\\
\textbf{With a more realistic model of a common cultural transmission bias, we may be able to better understand decision-making processes in humans, including life-changing choices such as occupation or a life partner.} 
Our model can be expanded in multiple ways: observing the effects of different bias functions, including errors in estimating the influence, combining factors of cost when copying from an influential role model (not all could afford to copy from the most popular role-model), and analyzing the differences when including several optimal values for the indicator trait (multiple preference traits in the population).

\subsection{Prestige in the literature}
So far we discussed prestige as a main bias in humans, and to some extent in non-human species. Here we further base our claims and present additional appearances of prestige in nature and in the literature. \\

\citet{animal_leadership} describe a manifestation of prestige in the form of leadership in animals. According to them, there are two main approaches to decision makings of groups in nature: leadership and consensus. Prey leaders would lead the pact when traveling, while other animal group leaders will decide on a nesting site or foraging patches. They found out that leadership is observed mostly when there is a profound social network in the group, and when there are individuals that present leadership behavior. Leaders would usually be high ranking members in the group, such as elders, individuals with many kin relations, or posses other dominant traits. When no individual posses such traits, or when the social network is lacking, a consensus is more likely to occur.
When a leader is present, they will have greater selection costs, such as higher risk for predation, being poisoned by unknown experimental patch, but also greater benefits. For example, given the route to the foraging site was successful, the leader and his closest followers would gain most of the food, unlike in a consensus, where the food would be shared more equally.
It appears leaders appear in simple organisms as well, like fish. In these organisms however, the leader would usually be the hungriest or the weakest, while the rest would prefer to follow, minimizing their costs. In baboons however, \citet{animal_leadership} describe many benefits for the closest associates of the dominant male, such as protection from predators. This is an instance of sexual-selection, where the leader, whose survival chances are lower, gains more sexual partners due to the benefits. ("The greater the risk, the greater the reward")
What they describe could be the origins of what we know today as prestige. In their paper, they show that in nature, when survival is the main concern at all times, the leaders wouldn't be chosen due to their superior abilities, but because they have the least to lose. When in said position of leadership, there are greater risks, but greater rewards to come with it. 
In humans, leadership also has its perks and costs. Leaders can make decisions that would benefit them and their closest followers the most, while still maintaining group cohesion. However, wrong decision making that would harm the group could result in harm (media, social status, even violent behavior of subjects on certain cases).
In our society it is less common to worry about mere survival, and so the prestigious positions, even though are not without risks and costs, are not as dangerous as for animals in nature. This may be the reason humans strive for the prestigious positions, as they may reap rewards greater than the risk and costs to achieve them. This is in complete contrast to animals, where the weakest/hungriest is driven to lead, compared to humans where leadership positions are mostly competitive.\\

\citet{dual_leadership} suggest a different view of leadership. They note that most discussions assume there is one type of leadership, as seen above, and so they differ in their definitions. \citet{dual_leadership} suggest a way to solve said contradiction by defining two types of leaderships: prestige-based and dominance-based.
They present classical views of leaderships by Confucius and Machiavelli. Confucius views leaders as role models who exercise influence through possessing superior knowledge, skills,
and (outstanding) personal qualities. This description is very similar to our indicator trait. By contrast, Machiavelli views leaders as rulers who exercise influence by imposing costs through (the threat of) punishment. They say that these two opposing views are both partially supported by the available evidence but each one on its own offers an incomplete view into the complex and dynamic processes of leadership.

Our current model doesn't reflect the model described in this article, but several adjustments could be made in order to match it. If we assume there's a correlation between trait value to a type of leadership (so in our binary model, one trait would be of prestige, and the other of dominance) we can implement their suggested model. For that we would need to add cost-benefit parameters, so the ones choosing prestige will be rewarded, but pay more, while the ones choosing dominance would pay less, but gain less benefits. It could be interesting to see the dynamics and relations between our model parameters and these cost-benefit parameters.\\

\citet{prestige_evolution} support said claim that there are two types of leadership, and also define the two as prestige based and dominance based leadership types.
By their definition, the latter is defined by acquiring social status by using aggression, intimidation and violence. It is also more common than prestige in non humans. Their definition of prestige is somewhat synonymous with ours. According to their manuscript, prestige is composed both of estimation in the eyes of people (our indicator/success trait) and commanding position in people's minds, i.e number of copiers people think they have, which they define as \textit{influence} (similar to our definition for influence).
In their paper, they show that prestige evolved from natural selection, as an efficient process to extract reproductive benefit from the flow of socially transmitted information. Simply put, prestige is a natural step where social learning exists, due to saving costs of individual learning.
It could be interesting in the future to expand our model using this idea: observing the copier trait of \textit{evaluation}, rather than only observing the evolution of the indicator trait copied.
\citet{prestige_evolution} suggest that the most skilled role-models will, on average, end up with most copiers. Their research, definitions and results, is consistent with ours.\\

So far we presented the theory behind prestige, and it's appearance in nature. The following will show the appearances of prestige biases in humans, and in recent times.

\citet{prestige_cultural_learning}, for example, tested the existence of prestige in young children. \citet{prestige_cultural_learning} report the first direct tests in children that suggest the existence of \textit{prestige bias}, a tendency to learn from individuals to whom others have preferentially attended, learned or deferred.
Their definition of prestige is similar to our \textit{influence bias}, and brings concrete proof of its existence and effects. Their study showed that the odds of 3-4 years-old children learning from an adult model to whom bystanders had previously preferentially attended for 10 seconds were over twice those of their learning from a model whom bystanders ignored.
In addition to this first study, they also discovered prestige effects are domain-sensitive. They saw that prestigious models were listened to by most when demonstrating artifact-use, but not as much as when presenting food preferences. It lead \citet{prestige_cultural_learning} to believe that when the trait is costly to learn individually, prestige will have a higher effect. It would be interesting to include costs in our model to try and observe these effects and their dynamics in the simulations of a larger population than this study.\\

\citet{fijian_social_bias} researched Fijian villages, looking for evidence of social learning biases and their origins. They mention that:
\begin{quote}
evolutionary theorists propose that natural selection has favored the emergence of psychological biases for learning from those individuals most likely to possess adaptive information.
\end{quote}
Their goal is to bridge from the laboratory to the field by examining if and how these biases emerge in a small-scale society.
During their research they found that:
\begin{quote}
Fijian villagers (ages 10 and up) are biased to learn from others perceived as more successful/knowledgeable, both within and across domains (prestige effects).
\end{quote}
Their research shows promising evidence for our prestige model, suggesting that copying from others who are \textit{perceived} as successful, rather than actually are successful. In their paper, they show that the social networks representing copier-role-model relationships are centralized, suggesting:
\begin{quote}
This degree of centralization is consistent with the prediction that people substantially share notions about who is a good cultural model (network centrality), but that individuals’ model selections are influenced by multiple factors.
\end{quote}
We see here support for both our indicator trait and our influence bias in their data.\\

Aside from children's learning biases and small villages in a relatively primitive population, we can see prestige in more advanced domains as well, like western medicine.\citet{medical_prestige} present a specific and important effect of prestige - its significance for medical specialties and diseases.
They examined literature from 1950 to 2005 regarding the effects of prestige on medicinal practices. They discovered that active, specialized, biomedical, and high-technological types of medicine on organs in the upper part of the bodies of young and middle-aged people were accorded high levels of prestige, while medicine with opposite characteristics had low levels of prestige. They have concluded that such differences in prestige bear consequences for actual priority setting in healthcare systems. They discovered that surgery counts as the most prestigious specialty, while psychiatry is the less prestigious. In addition, doctors tend to rank practices that require more time to master as more prestigious, while other procedures that are considered \textit{easier} are less prestigious. Simply put, they found that the advance in technology and research was in accordance with the prestige rankings.
This means that there may be very important practices that are neglected due to the prestige bias.\\

As we seen so far, prestige can explain many behaviors and evolution of cultural traits. It is a tool to cheaply estimate and acquire knowledge, which helps an individual to survive and breed. However, it is not always the case, and there could be negative repercussions to this bias, such as invasion of maladaptive traits.

\citet{best_of_k} mention that social learning not only takes the form of random copying of other individuals, but also involves learners’ choice of what to learn and from whom to learn. They suggest a best-of-K model where an individual samples k role-models and choose the one he deems most "successful". They mentioned that a previous mathematical analysis has shown that it may sometimes result in maladaptive cultural evolution when the payoffs associated with cultural variants vary stochastically. In such a case, learners may be selectively disfavored and in the long run replaced by unbiased learners, who simply copy someone chosen at random. They develop new mathematical models that are simpler and mathematically tractable. They found that best-of-k learning, unlike unbiased learning, can facilitate the invasion of an on average inferior variant that sometimes gives a very high payoff. Our model, which includes influence bias, is consistent with that claim. When a maladaptive trait is "piggybacking" a role-model with high influence, said trait could spread in the population, as mentioned. In addition, they show that best-of-k learning can be stable against invasion by unbiased learning if social learning is sometimes combined with individual learning. Our model is based on copying based learning only, but it could be interesting to combine it with individual learning and see how it affects the dynamics of the population.

We discussed prestige in depth, and provided several proofs for its existence in nature, humans, and even medicine. We saw it could aid invasion of maladaptive traits, or neglect of important medicinal specialities. But, it can also accelerate reversal of harmful traditions.
Harmful traditions can be child marriage, open defecation, and domestic violence, to name a few. 
\citet{harmful_traditions} suggest a mechanism called \textit{spillover}. By their definition, a spillover is when an intervention affect a large enough group in a target population, so that others not included in the intervention starts changing their behavior as well. In their research, they found that there are individuals who act as \textit{agents}, who are often looked upon, and therefore they are ideal targets for interventions. This is the same concept as our role-models, where a more prestigious individual will be copied more, therefore spreading his trait wider in the population. Their research support therefore in our assumption that there are social biases, conformist influence specifically. They also suggest a way to use this phenomena to change existing traditions in a population. It is very clear however, that just as it can be used to end harmful traditions, the same agents could be used for any negative way that comes to mind.\\

Up until now, we showed that cultural transmission is a process that manifests in many species, with emphasis on humans. We also displayed similarities between this process and genetic transmission, while presenting differences between them, specifically selective biases such as influence and prestige. We also presented examples of good and bad usages of such biases. All of these are mainly presented as a parallel process to the natural selection process in regards to physical anatomy, or at least have an indirect effect on it. \citet{collective_brains} offer a take on prestige as a factor of human physical evolution directly. They present a concept called \textit{cultural brains} - brains that evolved primarily for the acquisition of adaptive knowledge.
They build on the hypothesis of \citet{social_brains} that shows that larger, more complex brains can store and manage more information and in turn, this information can support the costs of a larger brain.
\citet{collective_brains} built a model that predicts a strong relationship between brain size and group size, because group size also provides access to more adaptive knowledge. They later present their \textit{cumulative cultural brain} hypothesis, an approach which proposes that human brains have evolved with an ability and proclivity for selective, high fidelity social learning. As part of this process, there are a variety of strategies and biases that have evolved to hone in on the most adaptive knowledge. These strategies and biases include direct and indirect cues of the popularity of cultural traits (e.g. success and prestige biases).
In short, they suggest that some of the reasons for the extreme increase in brain size in humans, are the ability to "cheaply" acquire adaptive knowledge, i.e transmission biases, such as prestige.
